### 2025-05-07 chapters/3-jets.tex
Completed problems and rough solutions

We have predicted the presence of collinear sprays of particles in scattering events the strong force and defined jets in order to capture their properties more precisely.
%
Now, we want to find a way to leverage jets to determine the fundamental properties of our universe.
%
In this thesis, we take the approach of studying the \textit{substructure} of jets.
%
The study of jet substructure has contributed to the deep connection between experimental and theoretical particle physics, with notable successes including precise determinations of the strong coupling constant \(\alpha_s\) and the on-going search for the precise value of the top quark mass \(m_t\) \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}, whose value dictates the stability of the \glslink{sm}{Standard Model} vacuum and therefore the fate of the universe as we know it.

@@

We will use the framework of renormalization in a very specific context:
%
what happens when we zoom in on jets.
%
The inner structure of jets is quite similar to the example of the snowflake -- much more than you might guess.
%
As in the case of snow, as we zoom out, simple fundamental building blocks (water molecules/high-energy partons) turn to fractals (snowflakes/the tree of partonic splittings) and then again to simple, but macroscopic objects (the snow bank/jets themselves).
%
See \Fig{renormalization}, which explores renormalization in the context of magnetism, snowflakes, and jets, and exemplifies how zooming into a system tends to reveal simple constituents, while zooming out of a system tends to reveal complex structures or disorder.

@@

Resummation is the technical segue we need to go from the conceptual developments of jets and renormalization to a quantitative study of jet substructure, and it is a necessary tool for removing the infinities and negative probabilities plaguing our the fixed order computations in \Sec{fixed-order-substructure}.
%
We perform simple resummed analyses in \Secs{ll-substructure-diagrams}{p2p-fragmentation} which leverage the recursive structure of the partonic cascade to obtain meaningful results at all orders in \(\alpha_s\) (unlike e.g. the analyses of \Sec{fixed-order-substructure}, which worked only to NLO).

@@

We develop our technical description of the partonic cascade in \Sec{everything-on-shell}.
%
For those who wish to bypass the technical introduction and quickly access perturbative computations, here is the main conclusion of \Sec{everything-on-shell}:
\begin{answer}
    The technology of splitting functions developed in the previous chapter can be applied recursively in the soft-collinear limit.
    %
    This leads to a model of jets in which partons split recursively with a phase space probability distribution given by
    \begin{align}
        \splitxsec{\ell}{ij}
        =
        \frac{\alpha_s}{\pi}
        P_{ij \leftarrow \ell}(z_i)\,\frac{\dd z_i \, \dd \theta}{\theta}
        \,,
    \end{align}
    where \(z_i = E_i / E_\ell\) and \(\theta = \theta_{ij}\), exactly as in \Eq{dsigmatilde_definition} of \Sec{splitting-functions}.
\end{answer}
In \Secs{ll-substructure-diagrams}{p2p-fragmentation}, use our model of the partonic cascade to perform explicit calculations that will allow us to perform realistic calculations of jet substructure in \Chaps{grooming}{ewocs}.
%
In the next section, we will discuss re-clustering and low-energy pollution in preparation for a more realistic analysis of experimentally-obtained jets.

@@

The model of the partonic cascade we have found above allows us to recursively calculate features of \(N\)-particle differential cross sections:
%
the probability to produce \(N\) particles is related to the probability to produce \(N-1\), and so on.
%
Framing our results in a probabilistic fashion paves the way to another idea, however:
%
we may use our insights to algorithmically \textit{produce} \(N\) particle configurations, given a configuration of some number \(N-M\) of original particles.
%
If we compute the differential cross section for a process with two outgoing, final-state particles, we can use our results to obtain the differential cross section with \textit{any} number of final states.
%
We split two particles into three, then three into four, and so on, holding always to the tenets of angular ordering and using the probability distribution
\begin{align}
    \label{eq:splitting-prob-ch3}
    \dsigmatilde_{ij \leftarrow \ell^*}
    =
    \frac{\alpha_s}{2\pi}
    P_{ij \leftarrow \ell}(z_i) \frac{\dd z_i \dd s}{s}
    =
    \frac{\alpha_s}{\pi}
    P_{ij \leftarrow \ell}(z_i) \frac{\dd z_i \, \dd \theta}{\theta}
    \,,
\end{align}
as in \Eq{dsigmatilde_definition} of \Sec{splitting-functions}, for each splitting.
%
This procedure allows us to simulate the energy flow of a particle collision event not in the formal language of quantum field theory, but as a \textbf{classical, Markovian branching process}.
%
Simulations of this nature -- called \vocab{parton showers} -- are employed in more complicated forms as some of the most important tools of particle physics.

@@

    Using the principles of strong angular ordering and the soft limit, we can write its cumulative probability distribution in the partonic cascade model as
    \begin{align}
        \label{eq:general-additive-cumulative-distribution}
        \Sigma_X(x)
        &=
        \delta(x)
        \\
        \notag
        &\quad
        +
        \sum_{i = n}^\infty
        \,\,
        \le(
        \int_0^R \frac{\dd\theta_1}{\theta_1}
        \int_0^1 \dd z_1
        \frac{\alpha_s}{\pi}
        p_i(z_1)
        \ri)
        \le(
        \int_0^{\theta_1} \frac{\dd\theta_2}{\theta_2}
        \int_0^1 \dd z_2
        \frac{\alpha_s}{\pi}
        p_i(z_2)
        \ri)
        \cdots
        \\
        &\qquad\qquad\qquad\qquad
        \Theta(x > \sum_{k = 1}^n X(z_k, \theta_k))
        \,,
    \end{align}
    where the delta function emerges because we assume \(X = 0\) when evaluated on a single particle, and the bounds of integration on the angular integrals are a result of strong angular ordering.
    %
    We may use \(p_i(z) \approx 2 C_{R_i} / z\) -- the soft limit for the emission of a gluon from the initiating parton -- for each emission, because the emission of a soft gluon of momentum fraction \(z\) does not change the flavor of a parton:
    %
    a quark emitting a gluon remains a quark, and similarly a gluon remains a gluon after the soft emission.

@@

    Therefore, we can write
    \begin{align}
        \label{eq:dysons-additive-cumulative-distribution}
        \Sigma_X(x)
        &=
        \delta(x)
        \,\,
        +
        \,\,
        \sum_{i = n}^\infty
        \frac{1}{n!}
        \,\,
        \prod_{k = 1}^n
        \le(
        \int_0^R \frac{\dd\theta_k}{\theta_k}
        \int_0^1 \dd z_k
        \frac{\alpha_s}{\pi}
        p_i(z_1)
        \ri)
        \,\,
        \Theta\le(x > \sum_{k = 1}^n X(z_k, \theta_k)\ri)
        \,,
    \end{align}
    providing us with an extremely compact form for the cumulative distributions of additive observables.

@@

\begin{proposition}{LL Distributions for Additive Observables}{ll-additive}
    We may describe the \glslink{accuracy}{LL} cumulative probability distribution \(\Sigma^{\rm(LL)}_X(x)\) of an additive observable \(X\) as the probability that \textit{no emission} contributes a value greater than \(x\) to \(X\), up to sub-leading behavior in \(\alpha_s\) and logarithms or powers of the observable.
    \begin{align}
        \label{eq:ll-additive}
        \Sigma^{\text{(LL)}}_X(x) \approx \mathbb{P}(X_i < x \text{ for all } i)
        \,.
    \end{align}

@@

In more mathematical language, let us consider an additive observable with contributions \(X(z_i, \theta_i) = X_i > 0\) from emissions \(i\).
%
At \glslink{accuracy}{LL}, we have asserted that there is a dominant contribution \(X = \max_i x_i\) such that \(X = \sum_i X_i \approx \max_i X_i\).
%
We may therefore approximate the cumulative probability distribution for \(X\), \(\Sigma_X(x) := \mathbb{P}(X = \sum_i X_i < x)\), by the probability that the value of \(X_i\) for any emission is less than the dominant value, which in turn is less than \(x\):%
\footnote{
    More generally, we can write \textit{without approximation} that \(\Theta(x - \sum_i X_i) \leq \prod_i \Theta(x - X_i)\), where equality emerges when we have at most a single non-zero \(x_j = X\).
    %
    Using that the cumulative distribution for \(X\) is the expectation value of the left-hand side, we may take the expectation value of both sides to obtain \(\Sigma_X(x) \leq \mathbb{P}(X_i < x \text{ for all } i)\), which provides a helpful upper bound for the full cumulative distribution of an additive observable.
}
\begin{align}
    \Sigma_X(x) \approx \mathbb{P}(X_i < x \text{ for all } i)
    \,.
\end{align}
%
This approximation holds up to corrections due to multiple emissions, which we address in \Prob{multiple-emissions}, that are sub-leading in \(\alpha_s\) and logarithms of the observable.
%
We should also note that typically, calculations applying the \glslink{accuracy}{LL} approximation will also drop terms that are suppressed by powers of \(X\) since these are sub-dominant to the leading logarithmic terms in observable distributions.
\end{proof}

@@

Using \Example{dyson-substructure} and \Prop{ll-additive}, and using the reduced splitting functions of \Sec{fixed-order-substructure}, the jet mass distribution is given at LL by
%
\begin{equation}
\label{eq:LL_mass}
\begin{aligned}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    &=
    \sum_{n = 0}^{\infty}\frac{1}{n!}
    \left(
    \frac{\alpha_s}{\pi}
    \int_0^R \frac{\dd\theta}{\theta}
    \int_0^{\frac{1}{2}} \dd z\,[\bar{p}_i(z)]_+^{[1/2]}
    \Theta\left(
    \frac{m^2}{Q^2} - z\theta^2
    \right)
    \right)^n
    \\
    &=
    \exp\left[
    -\frac{\alpha_s}{\pi}
    \int_0^R \frac{\dd\theta}{\theta}
    \int_0^{\frac{1}{2}} \dd z\,\bar{p}_i(z)
    \Theta\left(
    z \theta^2 - \frac{m^2}{Q^2}
    \right)
    \right]
    .
\end{aligned}
\end{equation}

Let us pursue this derivation graphically by using \glspl{substructure-diagram}.
%
By following the integral expressions of \Eq{LL_mass}, the LL cumulative distribution for the mass-squared of a jet is
\begin{equation}
\begin{aligned}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    &=
    \sum_{n = 0}^{\infty}\frac{1}{n!}
    \left(
    \begin{tikzpicture}[
    baseline={([yshift=-.8ex]current bounding box.center)},
    vertex/.style={anchor=base,
    circle,fill=black!25,minimum size=18pt,inner sep=2pt},
    scale=.3]
    \begin{axis}
    [xmin=0, xmax=5,
    ymin=0, ymax=5,
    axis line style = very thick,
    axis y line*=left,
    axis x line*=bottom,
    axis lines = middle,
    ticks=none]
    	\addplot[name path=f, domain=0:5,
        style=very thick, blue]
        {4.5-1.5*x};
        \path[name path=axis]
        (axis cs:0,5) -- (axis cs:5,5);
        \addplot [
            thick,
            color=blue,
            fill=blue,
            fill opacity=0.3
        ]
        fill between[
            of=f and axis
        ];
        \node [rotate=-55] at (axis cs:  3.2,  2.8)
        {${\scriptstyle \frac{m^2}{Q^2} > z\theta^2}$};
    \end{axis}
    \end{tikzpicture}
    \right)^n
    =
    \exp\left[
    ~-~
    \begin{tikzpicture}[
    baseline={([yshift=-.8ex]current bounding box.center)},
    vertex/.style={anchor=base,
    circle,fill=black!25,minimum size=18pt,inner sep=2pt},
    scale=.3]
    \begin{axis}
    [xmin=0, xmax=5,
    ymin=0, ymax=5,
    axis line style = very thick,
    axis y line*=left,
    axis x line*=bottom,
    axis lines = middle,
    ticks=none]
    	\addplot[name path=f, domain=0:5,
        style=very thick, blue]
        {4.5-1.5*x};
        \path[name path=axis]
        (axis cs:0,0) -- (axis cs:1,0);
        \addplot [
            thick,
            color=blue,
            fill=blue,
            fill opacity=0.3
        ]
        fill between[
            of=f and axis
        ];
        \node [rotate=-50] at (axis cs:  1.25,  1.25)
        {${\scriptstyle \frac{m^2}{Q^2} < z\theta^2}$};
    \end{axis}
    \end{tikzpicture}
    ~~
    \right]
    %\approx
    %\exp\left[~-\frac{C_R\alpha_s}{2 %\pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    ,
\end{aligned}
\end{equation}
where the first equality gives the probability that no emission in the jet is associated with an approximate mass-squared \(z\theta^2\) larger than \(m^2\) -- the argument of the cumulative distribution -- and the second equality follows from using the fact that \glspl{plus-fn} integrate to zero (see the diagrammatic identity of \Eq{plus-fn-diagram-identities}).
%
In the final equality, we see that the LL cumulative distribution is one over the exponent of the \gls{radiator} (the area of the \gls{vetoreg}).

Plugging in the important piece for the splitting function at leading logarithmic accuracy (i.e. in the soft limit), \(p(z) \approx 2 C_R / z\) for a jet formed by particles with \(SU(3)\) color representation \(R\), and putting back in factors of \(Q\), we find
\begin{align}
    \label{eq:mass-ll-cumulative}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    =
    \exp\left[~-\frac{C_R\alpha_{\rm S}}{2 \pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    \,,
\end{align}
%
with an associated probability distribution
\begin{align}
    \rho^{\rm(LL)}_{m^2}(m^2)
    =
    \frac{C_R \alpha_s}{\pi}
    \frac{1}{m^2}
    \log\le(\frac{Q^2}{m^2}\ri)
    \,
    \exp\left[~-\frac{C_R\alpha_{\rm S}}{2 \pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    \,.
\end{align}

@@

%\begin{exercise}
%    When expanded to \(\mathcal{O}(\alpha_s)\), there is a small discrepancy between the LL result for the cumulative mass-squared distribution above and the NLO result presented in \Sec{fixed-order-substructure}.

@@

\begin{exercise}
    The slopes of the diagonal line in the substructure diagram in the calculation of the jet mass distribution are \(-2\) since, for constant \(m^2\), \(z \sim 1/\theta^2\) and \(\log(1/z) = \text{const} - 2 \log(1/\theta)\).
    %
    What are the slopes of the diagonal lines involved in the computation of jet \glspl{angularity}?

    Represent the LL distribution of jet angularities with a \gls{substructure-diagram}, and show that it takes the approximate analytic form
    \begin{align}
        \Sigma^{(\text{LL})}(e_\varsigma)
        \approx
        \exp\le[
            -\frac{C_R\alpha_{\rm S}}{\varsigma \pi}\log^2(e_\varsigma)
        \ri]
        \,,
    \end{align}
    in the soft limit.
\end{exercise}

@@

\remark{}{
    To include running coupling effects and work towards an MLL description of additive jet observables, we notice that the splittings of perturbative QCD should be governed by an \textit{effective coupling} \(\alpha_s(\kappa)\), where \(\kappa = \kappa(z,\theta)\) is a scale associated with the splitting.
    %
    The result would involve integrals containing terms of the form \(\int \dd z \, \dd \theta \, \alpha_s(\kappa) \, p(z) / \theta\).
    %
    The precise value we should choose for \(\kappa\) in order to account for higher-order effects is not obvious, but \Reff{Amati:1980ch} showed that \(\kappa\) should be given by the transverse momentum of the splitting in order to minimize the effects of large logarithms in perturbative calculations:
    \begin{align}
        \kappa \approx Q z(1-z)\sin\theta \approx Q z \theta
        \,.
    \end{align}
    %
    In particular, QCD correlations and cross sections that use \(\alpha_s\) evaluated at a scale \(\kappa\) generically include logarithms of the form \(\log(p_{\perp} / \kappa)\).
    %
    To make these logarithms small, one may choose \(\kappa \sim p_\perp\), which then requires that we evaluate the strong coupling at the scale \(p_\perp\) as well.
    %
    This prescription provides a straightforward procedure for understanding higher-order effects with \glspl{substructure-diagram}, and can be extended to higher perturbative accuracy with the Catani-Marchesini-Webber (CMW) prescription \cite{Catani:1990rr}, which achieves two-loop accuracy for the running of \(\alpha_s\), and to qualitatively include non-perturbative effects by \textit{freezing} the coupling, as discussed in \App{soft-distortions}.
    %
    As a reminder, the one-loop running of \(\alpha_s\) is captured by
    \begin{align}
        \alpha^{\rm 1-loop}_s(\kappa)
        =
        \frac{\alpha_s(Q)}{1 + \alpha_s(Q) \beta_0 \log(\frac{\kappa}{Q})/2\pi}
        ,
    \end{align}
    where \(\beta_0 = \frac{11}{3}C_A - \frac{4}{3} T_f n_f\) is a coefficient appearing in the QCD beta function: \(\beta(\alpha_s) = -\beta_0 \alpha_s^2/2\pi + \mathcal{O}(\alpha_s^3)\).
}

>>>>

We have predicted the presence of collinear sprays of particles in scattering events the strong force and defined jets in order to capture their properties more precisely.
%
Now, we want to find a way to leverage jets to determine the fundamental properties of our universe.
%
In this thesis, we take the approach of studying the \textit{substructure} of jets.
%
The study of jet substructure has contributed to the deep connection between experimental and theoretical particle physics, with notable successes including precise determinations of the strong coupling constant \(\alphas\) and the on-going search for the precise value of the top quark mass \(m_t\) \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}, whose value dictates the stability of the \glslink{sm}{Standard Model} vacuum and therefore the fate of the universe as we know it \cite{Alekhin:2012py}.

@@

We will use the framework of renormalization in a very specific context:
%
what happens when we zoom in on jets.
%
The inner structure of jets is quite similar to the example of the snowflake -- much more than you might guess.
%
As in the case of snow, as we zoom out, simple fundamental building blocks (water molecules/high-energy partons) turn to fractals (snowflakes/the tree of partonic splittings) and then again to simple, but macroscopic objects (the snow bank/jets themselves).

@@

Resummation is the technical segue we need to go from the conceptual developments of jets and renormalization to a quantitative study of jet substructure, and it is a necessary tool for removing the infinities and negative probabilities plaguing our the fixed order computations in \Sec{fixed-order-substructure}.
%
We perform simple resummed analyses in \Secs{ll-substructure-diagrams}{p2p-fragmentation} which leverage the recursive structure of the partonic cascade to obtain meaningful results at all orders in \(\alphas\) (unlike e.g. the analyses of \Sec{fixed-order-substructure}, which worked only to NLO).

@@

We develop our technical description of the partonic cascade in \Sec{everything-on-shell}.
%
For those who wish to bypass the technical introduction and quickly access perturbative computations, here is the main conclusion of \Sec{everything-on-shell}:
\begin{answer}
    The technology of splitting functions developed in the previous chapter can be applied recursively in the soft-collinear limit.
    %
    This leads to a model of jets in which partons split recursively with a phase space probability distribution given by
    \begin{align}
        \splitxsec{\ell}{ij}
        =
        \frac{\alphas}{\pi}
        P_{ij \leftarrow \ell}(z_i)\,\frac{\dd z_i \, \dd \theta}{\theta}
        \,,
    \end{align}
    where \(z_i = E_i / E_\ell\) and \(\theta = \theta_{ij}\), exactly as in \Eq{dsigmatilde_definition} of \Sec{splitting-functions}.
\end{answer}
In \Secs{ll-substructure-diagrams}{p2p-fragmentation}, use our model of the partonic cascade to perform explicit calculations that will allow us to perform realistic calculations of jet substructure in \Chaps{grooming}{ewocs}.
%
In the next section, we will discuss re-clustering and low-energy pollution in preparation for a more realistic analysis of experimentally-obtained jets.

@@

The model of the partonic cascade we have found above allows us to recursively calculate features of \(N\)-particle differential cross sections:
%
the probability to produce \(N\) particles is related to the probability to produce \(N-1\), and so on.
%
Framing our results in a probabilistic fashion paves the way to another idea, however:
%
we may use our insights to algorithmically \textit{produce} \(N\) particle configurations, given a configuration of some number \(N-M\) of original particles.
%
If we compute the differential cross section for a process with two outgoing, final-state particles, we can use our results to obtain the differential cross section with \textit{any} number of final states.
%
We split two particles into three, then three into four, and so on, holding always to the tenets of angular ordering and using the probability distribution
\begin{align}
    \label{eq:splitting-prob-ch3}
    \dsigmatilde_{ij \leftarrow \ell^*}
    =
    \frac{\alphas}{2\pi}
    P_{ij \leftarrow \ell}(z_i) \frac{\dd z_i \dd s}{s}
    =
    \frac{\alphas}{\pi}
    P_{ij \leftarrow \ell}(z_i) \frac{\dd z_i \, \dd \theta}{\theta}
    \,,
\end{align}
as in \Eq{dsigmatilde_definition} of \Sec{splitting-functions}, for each splitting.
%
This procedure allows us to simulate the energy flow of a particle collision event not in the formal language of quantum field theory, but as a \textbf{classical, Markovian branching process}.
%
Simulations of this nature -- called \vocab{parton showers} -- are employed in more complicated forms as some of the most important tools of particle physics.

@@

    Using the principles of strong angular ordering and the soft limit, we can write its cumulative probability distribution in the partonic cascade model as
    \begin{align}
        \label{eq:general-additive-cumulative-distribution}
        \Sigma_X(x)
        &=
        \delta(x)
        \\
        \notag
        &\quad
        +
        \sum_{i = n}^\infty
        \,\,
        \le(
        \int_0^R \frac{\dd\theta_1}{\theta_1}
        \int_0^1 \dd z_1
        \frac{\alphas}{\pi}
        p_i(z_1)
        \ri)
        \le(
        \int_0^{\theta_1} \frac{\dd\theta_2}{\theta_2}
        \int_0^1 \dd z_2
        \frac{\alphas}{\pi}
        p_i(z_2)
        \ri)
        \cdots
        \\
        &\qquad\qquad\qquad\qquad
        \Theta(x > \sum_{k = 1}^n X(z_k, \theta_k))
        \,,
    \end{align}
    where the delta function emerges because we assume \(X = 0\) when evaluated on a single particle, and the bounds of integration on the angular integrals are a result of strong angular ordering.
    %
    We may use \(p_i(z) \approx 2 C_{R_i} / z\) -- the soft limit for the emission of a gluon from the initiating parton -- for each emission, because the emission of a soft gluon of momentum fraction \(z\) does not change the flavor of a parton:
    %
    a quark emitting a gluon remains a quark, and similarly a gluon remains a gluon after the soft emission.

@@

    Therefore, we can write
    \begin{align}
        \label{eq:dysons-additive-cumulative-distribution}
        \Sigma_X(x)
        &=
        \delta(x)
        \,\,
        +
        \,\,
        \sum_{i = n}^\infty
        \frac{1}{n!}
        \,\,
        \prod_{k = 1}^n
        \le(
        \int_0^R \frac{\dd\theta_k}{\theta_k}
        \int_0^1 \dd z_k
        \frac{\alphas}{\pi}
        p_i(z_1)
        \ri)
        \,\,
        \Theta\le(x > \sum_{k = 1}^n X(z_k, \theta_k)\ri)
        \,,
    \end{align}
    providing us with an extremely compact form for the cumulative distributions of additive observables.

@@

\begin{proposition}{LL Distributions for Additive Observables}{ll-additive}
    We may describe the \glslink{accuracy}{LL} cumulative probability distribution \(\Sigma^{\rm(LL)}_X(x)\) of an additive observable \(X\) as the probability that \textit{no emission} contributes a value greater than \(x\) to \(X\), up to sub-leading behavior in \(\alphas\) and logarithms or powers of the observable.
    \begin{align}
        \label{eq:ll-additive}
        \Sigma^{\text{(LL)}}_X(x) \approx \mathbb{P}(X_i < x \text{ for all } i)
        \,.
    \end{align}

@@

In more mathematical language, let us consider an additive observable with contributions \(X(z_i, \theta_i) = X_i > 0\) from emissions \(i\).
%
At \glslink{accuracy}{LL}, we have asserted that there is a dominant contribution \(X = \max_i x_i\) such that \(X = \sum_i X_i \approx \max_i X_i\).
%
We may therefore approximate the cumulative probability distribution for \(X\), \(\Sigma_X(x) := \mathbb{P}(X = \sum_i X_i < x)\), by the probability that the value of \(X_i\) for any emission is less than the dominant value, which in turn is less than \(x\):%
\footnote{
    More generally, we can write \textit{without approximation} that \(\Theta(x - \sum_i X_i) \leq \prod_i \Theta(x - X_i)\), where equality emerges when we have at most a single non-zero \(x_j = X\).
    %
    Using that the cumulative distribution for \(X\) is the expectation value of the left-hand side, we may take the expectation value of both sides to obtain \(\Sigma_X(x) \leq \mathbb{P}(X_i < x \text{ for all } i)\), which provides a helpful upper bound for the full cumulative distribution of an additive observable.
}
\begin{align}
    \Sigma_X(x) \approx \mathbb{P}(X_i < x \text{ for all } i)
    \,.
\end{align}
%
This approximation holds up to corrections due to multiple emissions, which we address in \Prob{multiple-emissions}, that are sub-leading in \(\alphas\) and logarithms of the observable.
%
We should also note that typically, calculations applying the \glslink{accuracy}{LL} approximation will also drop terms that are suppressed by powers of \(X\) since these are sub-dominant to the leading logarithmic terms in observable distributions.
\end{proof}

@@

Using \Example{dyson-substructure} and \Prop{ll-additive}, and using the reduced splitting functions of \Sec{fixed-order-substructure}, the jet mass distribution is given at LL by
%
\begin{equation}
\label{eq:LL_mass}
\begin{aligned}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    &=
    \sum_{n = 0}^{\infty}\frac{1}{n!}
    \left(
    \frac{\alphas}{\pi}
    \int_0^R \frac{\dd\theta}{\theta}
    \int_0^{\frac{1}{2}} \dd z\,[\bar{p}_i(z)]_+^{[1/2]}
    \Theta\left(
    \frac{m^2}{Q^2} - z\theta^2
    \right)
    \right)^n
    \\
    &=
    \exp\left[
    -\frac{\alphas}{\pi}
    \int_0^R \frac{\dd\theta}{\theta}
    \int_0^{\frac{1}{2}} \dd z\,\bar{p}_i(z)
    \Theta\left(
    z \theta^2 - \frac{m^2}{Q^2}
    \right)
    \right]
    .
\end{aligned}
\end{equation}

Let us pursue this derivation graphically by using \glspl{substructure-diagram}.
%
By following the integral expressions of \Eq{LL_mass}, the LL cumulative distribution for the mass-squared of a jet is
\begin{equation}
\begin{aligned}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    &=
    \sum_{n = 0}^{\infty}\frac{1}{n!}
    \left(
    \begin{tikzpicture}[
    baseline={([yshift=-.8ex]current bounding box.center)},
    vertex/.style={anchor=base,
    circle,fill=black!25,minimum size=18pt,inner sep=2pt},
    scale=.3]
    \begin{axis}
    [xmin=0, xmax=5,
    ymin=0, ymax=5,
    axis line style = very thick,
    axis y line*=left,
    axis x line*=bottom,
    axis lines = middle,
    ticks=none]
    	\addplot[name path=f, domain=0:5,
        style=very thick, blue]
        {4.5-1.5*x};
        \path[name path=axis]
        (axis cs:0,5) -- (axis cs:5,5);
        \addplot [
            thick,
            color=blue,
            fill=blue,
            fill opacity=0.3
        ]
        fill between[
            of=f and axis
        ];
        \node [rotate=-55] at (axis cs:  3.2,  2.8)
        {${\scriptstyle \frac{m^2}{Q^2} > z\theta^2}$};
    \end{axis}
    \end{tikzpicture}
    \right)^n
    =
    \exp\left[
    ~-~
    \begin{tikzpicture}[
    baseline={([yshift=-.8ex]current bounding box.center)},
    vertex/.style={anchor=base,
    circle,fill=black!25,minimum size=18pt,inner sep=2pt},
    scale=.3]
    \begin{axis}
    [xmin=0, xmax=5,
    ymin=0, ymax=5,
    axis line style = very thick,
    axis y line*=left,
    axis x line*=bottom,
    axis lines = middle,
    ticks=none]
    	\addplot[name path=f, domain=0:5,
        style=very thick, blue]
        {4.5-1.5*x};
        \path[name path=axis]
        (axis cs:0,0) -- (axis cs:1,0);
        \addplot [
            thick,
            color=blue,
            fill=blue,
            fill opacity=0.3
        ]
        fill between[
            of=f and axis
        ];
        \node [rotate=-50] at (axis cs:  1.25,  1.25)
        {${\scriptstyle \frac{m^2}{Q^2} < z\theta^2}$};
    \end{axis}
    \end{tikzpicture}
    ~~
    \right]
    %\approx
    %\exp\left[~-\frac{C_R\alphas}{2 %\pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    ,
\end{aligned}
\end{equation}
where the first equality gives the probability that no emission in the jet is associated with an approximate mass-squared \(z\theta^2\) larger than \(m^2\) -- the argument of the cumulative distribution -- and the second equality follows from using the fact that \glspl{plus-fn} integrate to zero (see the diagrammatic identity of \Eq{plus-fn-diagram-identities}).
%
In the final equality, we see that the LL cumulative distribution is one over the exponent of the \gls{radiator} (the area of the \gls{vetoreg}).

Plugging in the important piece for the splitting function at leading logarithmic accuracy (i.e. in the soft limit), \(p(z) \approx 2 C_R / z\) for a jet formed by particles with \(SU(3)\) color representation \(R\), and putting back in factors of \(Q\), we find
\begin{align}
    \label{eq:mass-ll-cumulative}
    \Sigma^{\rm(LL)}_{m^2}(m^2)
    =
    \exp\left[~-\frac{C_R\alphas}{2 \pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    \,,
\end{align}
%
with an associated probability distribution
\begin{align}
    \rho^{\rm(LL)}_{m^2}(m^2)
    =
    \frac{C_R \alphas}{\pi}
    \frac{1}{m^2}
    \log\le(\frac{Q^2}{m^2}\ri)
    \,
    \exp\left[~-\frac{C_R\alphas}{2 \pi}\log^2\left(\frac{m^2}{Q^2}\right)\right]
    \,.
\end{align}

@@

%\begin{exercise}
%    When expanded to \(\mathcal{O}(\alphas)\), there is a small discrepancy between the LL result for the cumulative mass-squared distribution above and the NLO result presented in \Sec{fixed-order-substructure}.

@@

\begin{exercise}
    \label{ex:ll-angularity-radiator}
    The slopes of the diagonal line in the substructure diagram in the calculation of the jet mass distribution are \(-2\) since, for constant \(m^2\), \(z \sim 1/\theta^2\) and \(\log(1/z) = \text{const} - 2 \log(1/\theta)\).
    %
    What are the slopes of the diagonal lines involved in the computation of jet \glspl{angularity}?

    Represent the LL distribution of jet angularities with a \gls{substructure-diagram}, and show that it takes the approximate analytic form
    \begin{align}
        \Sigma^{(\text{LL})}(e_\varsigma)
        \approx
        \exp\le[
            -\frac{C_R\alphas}{\varsigma \pi}\log^2(e_\varsigma)
        \ri]
        \,,
    \end{align}
    in the soft limit.
\end{exercise}

@@

\remark{}{
    To include running coupling effects and work towards an MLL description of additive jet observables, we notice that the splittings of perturbative QCD should be governed by an \textit{effective coupling} \(\alphas(\kappa)\), where \(\kappa = \kappa(z,\theta)\) is a scale associated with the splitting.
    %
    The result would involve integrals containing terms of the form \(\int \dd z \, \dd \theta \, \alphas(\kappa) \, p(z) / \theta\).
    %
    The precise value we should choose for \(\kappa\) in order to account for higher-order effects is not obvious, but \Reff{Amati:1980ch} showed that \(\kappa\) should be given by the transverse momentum of the splitting in order to minimize the effects of large logarithms in perturbative calculations:
    \begin{align}
        \kappa \approx Q z(1-z)\sin\theta \approx Q z \theta
        \,.
    \end{align}
    %
    In particular, QCD correlations and cross sections that use \(\alphas\) evaluated at a scale \(\kappa\) generically include logarithms of the form \(\log(p_{\perp} / \kappa)\).
    %
    To make these logarithms small, one may choose \(\kappa \sim p_\perp\), which then requires that we evaluate the strong coupling at the scale \(p_\perp\) as well.
    %
    This prescription provides a straightforward procedure for understanding higher-order effects with \glspl{substructure-diagram}, and can be extended to higher perturbative accuracy with the Catani-Marchesini-Webber (CMW) prescription \cite{Catani:1990rr}, which achieves two-loop accuracy for the running of \(\alphas\), and to qualitatively include non-perturbative effects by \textit{freezing} the coupling, as discussed in \App{soft-distortions}.
    %
    As a reminder, the one-loop running of \(\alphas\) is captured by
    \begin{align}
        \alpha^{\rm 1-loop}_s(\kappa)
        =
        \frac{\alphas(Q)}{1 + \alphas(Q) \beta_0 \log(\frac{\kappa}{Q})/2\pi}
        ,
    \end{align}
    where \(\beta_0 = \frac{11}{3}C_A - \frac{4}{3} T_f n_f\) is a coefficient appearing in the QCD beta function: \(\beta(\alphas) = -\beta_0 \alphas^2/2\pi + \mathcal{O}(\alphas^3)\).
}

To include more detail about the effects of multiple emissions, and achieve a full MLL description of additive jet obervables, we will need to use the full expression \(X = \sum_i X_i\) rather than the LL approximation \(X \approx \max_i X_i\) used above.
%
Concretely, this means that the effects of multiple emissions for an additive observable can be computed via the infinite-product integral
\begin{align}
    \label{eq:series-multiple-emissions}
    \Sigma_X(x)
    &=
    \sum_{n = 0}^{\infty}
    \frac{1}{n!}
    \prod_{k = 1}^n
    \left(
        \int_0^R \frac{\dd\theta_k}{\theta_k}
        \int_0^{\frac{1}{2}} \dd z_k
        \,\,
        \frac{\alphas}{\pi}
        \,\,
        [\bar{p}_i(z_k)]_+^{[1/2]}
    \right)^n
    \Theta\left(
        \frac{m^2}{Q^2} - \sum_k X(z_k, \theta_k)
    \right)
    \,.
\end{align}
%
At face value, this seems much more difficult to evaluate than the LL integral for the radiator of \(X\).
%
It is not at all obvious whether the contributions of multiple emissions exposed by \Eq{series-multiple-emissions} will even lead to an exponential form for the cumulative distribution of \(X\) --
indeed, \textit{they don't}.
